---
title: "Vehicle Towing in Buffalo"
author: Kaiser Mahmood
subtitle: Analysis in R
---

# Introduction

Towing of vehicles is a huge problem and there may be a lot of factor that could influence towing. In this project I tried to explore the towing data in buffalo to figure out any potential spatial relation of towing with location, and roadway and demographic characteristics.

# Materials and methods

You can do numbers like this:

1. Getting associated data
2. Data cleaning
3. Getting Spatial map


Load any required packages in a code chunk (you may need to install some packages):

```{r, message=F, warning=F}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggmap)
library(sf)
library(sp)
library(kableExtra)
```

## Download and clean all required data
```{r, message=F, warning=F}
data=read_csv("https://data.buffalony.gov/api/views/5c88-nfii/rows.csv?accessType=DOWNLOAD")
```


```{r}
plot1 <- ggplot(data,aes(LONGITUDE,LATITUDE,col=`TOW DESCRIPTION`))+geom_point()
plot1
```

<!--```{r}
library(tidyverse)
library(ggmap)
?get_map
buffalo <- c(lon=-78,lat=44)
buffalo_map <- get_map(location = buffalo,zoom=5)
plot2 <- ggmap(buffalo_map)+geom_point(d,aes(LONGITUDE,LATITUDE))
plot2
```
-->

Add any additional processing steps here.

# Results


# Conclusions



# References


