---
title: "Vehicle Towing in Buffalo"
author: Kaiser Mahmood
subtitle: Analysis in R
output:
  html_document:
    code_folding: show
---

# **Introduction**

Towing of vehicles is a huge problem and there may be a lot of factor that could influence towing. In this project I tried to explore the towing data in buffalo to figure out any potential spatial relation of towing with location, and roadway and demographic characteristics.

# **Materials and methods**

You can do numbers like this:S

1. Getting associated data
2. Data cleaning
3. Getting interactive map


Load any required packages in a code chunk (you may need to install some packages):

```{r Loading Library, results = 'hide', message=F, warning=F, echo=TRUE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(lubridate)
library(leaflet)
library(ggmap)
library(sf)
library(sp)
library(kableExtra)
library(htmlwidgets)
library(widgetframe)
library(rnoaa)
library(xts)
library(dygraphs)
```

## **Download and clean all required data**
```{r data cleaning, results = 'hide', message = F, warning = F,echo=TRUE}
data=read_csv("https://data.buffalony.gov/api/views/5c88-nfii/rows.csv?accessType=DOWNLOAD")

data2 <- separate(data, col = `TOW DATE`, into = c("Month","Day", "Year"), sep = "/") %>%
  select(Month, Day, Year,`TOW DESCRIPTION`, LATITUDE, LONGITUDE, `POLICE DISTRICT`) %>%
  filter(LONGITUDE!="NA"|LATITUDE!="NA")

data2$Month <- as.numeric(data2$Month)
data2$Year <- as.numeric(data2$Year)
data2$Day <- as.numeric(data2$Day)
data2$Date <- as.Date(with(data2, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")

Tow_count <- data2 %>% group_by(Date) %>% summarize(count=n())
tow_xts <- xts(Tow_count$count,order.by = Tow_count$Date)

#data2 %>% 
#  slice(1:10) %>% #show only 1:n rows
#  kable(digits=2,align="c")%>% #make table and round to two digits
#  kable_styling(bootstrap_options = 
#                  c("striped", "hover", "condensed", "responsive")) #apply other formatting

tow_reason <- data2 %>% group_by(`TOW DESCRIPTION`) %>% summarize(count=n()) %>% rename("Number of Tows"=count)
Police_district <- data2 %>% group_by(`POLICE DISTRICT`) %>% summarize(count=n()) %>% rename("Number of Tows"=count)
year_count <- data2 %>% group_by(Year) %>% summarize(count=n()) %>% rename("Number of Tows"=count)
month_count <- data2 %>% group_by(Month) %>% summarize(count=n()) %>% rename("Number of Tows"=count)
accident_count <- data2 %>% filter(`TOW DESCRIPTION`=="ACCIDENT")
#month_year <- data2 %>% group_by(Month, Year) %>% summarize(count=n())

data2$Month <- as.factor(data2$Month)
levels(data2$Month) <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "July", "Aug", "Sep", "Oct","Nov", "Dec")
month_year <- data2 %>% group_by(Month, Year) %>% summarize(count=n()) %>% rename("Number of Tows"=count)

# needs asthetic
plot1 <- ggplot(year_count,aes(Year, `Number of Tows`)) + geom_col() 
plot2 <- ggplot(month_year,aes(Month, `Number of Tows`)) + geom_col()
plot3 <- ggplot(month_year,aes(Month,`Number of Tows`)) + geom_col() + facet_wrap(~Year)
```



# **Results**

## *Number of Tow in different Police district*
```{r Police district, message = F, warning = F,echo=FALSE}
Police_district %>% 
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive"))
```

## *Reasons for Towing*
```{r Towing reason, message=F, warning=F,echo=FALSE}
tow_reason %>% 
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive")) #apply other formatting
```

## *Yearly Towing Counts*
```{r yearly variation plot,  message = F, warning = F,echo=FALSE}
plot1
```

## *Monthly Variation in Tow Counts*
```{r monthy variation plot, message = F, warning = F,echo=FALSE}
plot2
plot3
```

## *Time series Plot of Daily Towing count in Buffalo*
```{r time series Plot, message = F, warning = F,echo=TRUE}
dygraph(tow_xts,main="Daily Tow count in Buffalo, NY")%>%
  dyRangeSelector(dateWindow = c("2007-01-01", "2019-03-26"))
```
*Please change range of time in the bottom to get more specific time data*



## *Showing Tow in interactive Map*
```{r interactive map, message = F, warning = F,echo=TRUE}
new_data <- data %>% filter(LONGITUDE!="NA"|LATITUDE!="NA")
all_tow <- leaflet(new_data) %>% 
   addTiles() %>% 
   setView( lng = -78.9, lat = 42.9, zoom = 11 ) %>% 
   addCircleMarkers(lng = ~LONGITUDE,
                    lat = ~LATITUDE,
                    clusterOptions = markerClusterOptions())%>%
   #addProviderTiles("Esri.WorldImagery")
   addProviderTiles("Esri.WorldTopoMap")
   #addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")
all_tow
```
*Please zoom in/out to see the location of towing*



## *Towing related to Accidents*
```{r interactive map accident, message = F, warning = F,echo=TRUE}
accident <- accident_count %>% filter(LONGITUDE!="NA"|LATITUDE!="NA")
accident_tow <- leaflet(accident) %>% 
   addTiles() %>% 
   setView( lng = -78.9, lat = 42.9, zoom = 11 ) %>% 
   addCircleMarkers(lng = ~LONGITUDE,
                    lat = ~LATITUDE,
                    clusterOptions = markerClusterOptions())%>%
   #addProviderTiles("Esri.WorldImagery")
   addProviderTiles("Esri.WorldTopoMap")
   #addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")
accident_tow
```
*Please zoom in/out to see the location of towing related to accident*



# **Conclusions**
A lot of the Tows are happening because of the accidents. If we look at the location of the tows related to accidents we can see that the accidents are occuring mostly at the intersection.Also the number of tows are increasing over the year.We can see from the table that police district D and E have the most number of tows.
For seasonal variation we can look at the histogram of tows with respect to the months.We can see from the histogram that january has a lot of tow occurance and then there is a decrease and then another increase in around july to august and then again decreasing gradually. December-january are the start of snow season. That's why I think there in an increase in the number of tows. Then people might get used to snow and the number of tow decrease. July, august are the time of summer and fall and people usually go out and make more recreational trips around this time which might be the reason for increase in trips.


# **References**

1. https://data.buffalony.gov/Transportation/Tows/5c88-nfii
2. DataCamp course list:
